### 2016-12-05 Poster/poster.tex
[minor] edit template file with current content

\documentclass[final,hyperref={pdfpagelabels=false}]{beamer}

\usepackage[orientation=portrait,size=a0,scale=1.4]{beamerposter}
\usetheme{I6pd2}
\usepackage[english]{babel}
\usepackage{amsmath,amsthm,amssymb,latexsym}
\usepackage{booktabs}
\usepackage{caption}
\usepackage{tcolorbox}
\usepackage{xcolor}
\graphicspath{{Images/}}
\usepackage[tight,footnotesize]{subfigure}

\definecolor{Blue}{HTML}{000080}

\usecaptiontemplate{\small\structure{\insertcaptionname~\insertcaptionnumber: }\insertcaption}

\title{Bayesian Nonparametric Motor-skill Representations \\ for Efficient Learning of Robotic Clothing Assistance}
\author{Nishanth Koganti$^{1,2}$, Tomoya Tamei$^1$, Kazushi Ikeda$^1$, Tomohiro Shibata$^2$}
\institute{$^1$Nara Institute of Science and Technology, Japan~~$^2$Kyushu Institute of Technology, Japan}

\newcommand{\leftfoot}{Neural Information Processing Systems 2016}
\newcommand{\rightfoot}{Workshop on Practical Bayesian Nonparametrics}

\begin{document}

\addtobeamertemplate{block end}{}{\vspace*{0.5ex}}
\addtobeamertemplate{block alerted end}{}{\vspace*{2ex}}

\begin{frame}[t]

\begin{columns}[t]
  \begin{column}{.02\linewidth}\end{column}

  \begin{column}{0.47\linewidth}
    \begin{block}{Introduction}
      \begin{center}
        \textbf{Abstract}: Motor-skill learning for complex robotic tasks is a challenging problem due to the high task variability. Robotic clothing assistance is one such challenging problem that can greatly improve the quality-of-life for the elderly and disabled. In this study, we propose a data-efficient representation to encode task specific motor-skills of the robot using Bayesian nonparametric latent space learning. We implement our framework in a practical setting with a dual-arm robot performing clothing assistance tasks. We demonstrate that performing motor-skills learning in such task specific latent spaces outperforms learning in the high-dimensional joint configuration space of the robot. Furthermore, our framework can also be used as a tool for learning from demonstration to impart novel skills to the robot.
      \end{center}
    \end{block}
  \end{column}

  \begin{column}{0.47\linewidth}
    \begin{block}{Problem Decription}

    \end{block}
  \end{column}

  \begin{column}{.02\linewidth}\end{column}
\end{columns}

\begin{columns}[t]
  \begin{column}{0.02\linewidth}\end{column}

  \begin{column}{0.96\linewidth}
    \begin{alertblock}{Motor-skill Representations in Latent Spaces}
      \begin{columns}[t]
        \begin{column}{0.24\linewidth}
          \centering \textbf{Learning from Demonstration}

        \end{column}

        \begin{column}{0.48\linewidth}
          \centering \textbf{Proposed Framework}

        \end{column}

        \begin{column}{0.24\linewidth}
          \centering \textbf{Reinforcement Learning Formulation}

        \end{column}
      \end{columns}
    \end{alertblock}
  \end{column}

  \begin{column}{0.02\linewidth}\end{column}
\end{columns}

\begin{alertblock}{Experimental Results}

\begin{columns}[t]

\begin{column}{0.49\linewidth}

\centering \textbf{}\\
\vspace{3mm}
\centering Robustness evaluated by comparing rewards obtained in \emph{Cartesian} and \emph{Topological} space.

\begin{itemize}
\item Reward function in Cartesian Space:
\begin{equation}
r(o_t) = - \sqrt{(o_t - o_{target})^T \Sigma_o^{-1} (o_t - o_{target})}
\end{equation}
\item Reward function in Topology Space:
\begin{equation}
r(s_t) = - \sqrt{(s_t - s_{target})^T \Sigma_s^{-1} (s_t - s_{target})}
\end{equation}
\end{itemize}

\begin{columns}[t]

\begin{column}{0.49\linewidth}

\centering Simulation Experiment
\begin{figure}
\centering
\includegraphics[width=\textwidth]{toptest.pdf}
\end{figure}

\end{column}

\begin{column}{0.49\linewidth}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{reward.png}
\end{figure}

\end{column}

\end{columns}

\end{column}

\begin{column}{0.49\linewidth}

\centering \textbf{Accuracy of Proposed Method}

\begin{itemize}
\item Accuracy of proposed method evaluated by comparing our estimates with the ground truth.
\vspace{3mm}
\item \textbf{Ground Truth}: Motion Capture System with 7 IR Cameras, 10 IR Markers used on T-shirt collar and 1 IR Marker on T-shirt sleeve.
\end{itemize}

\begin{columns}[t]

\begin{column}{0.59\linewidth}

\begin{figure}
\centering
\includegraphics[width=\textwidth]{realtimeresults.pdf}
\end{figure}

\end{column}

\begin{column}{0.39\linewidth}

\begin{table}
\centering
\begin{center}
\small
\begin{tabular}{ccc}
\toprule
Parameter & Pearson Corr & RMS Error\\[0.2cm]
\midrule
$s_1$ topology & 0.9364 & 0.5955\\
$s_2$ topology & 0.8857 & 0.1247\\
$s_3$ topology & 0.9889 & 0.3529\\
\bottomrule
\end{tabular}
\end{center}
\end{table}

\vspace{4mm}

\begin{itemize}
\item Proposed method provides reliable estimates in real-time (28-30 fps).
\vspace{3mm}
\item However, there are jitters due to inherent noise in depth sensor.
\end{itemize}

\end{column}

\end{columns}

\end{column}

\end{columns}

\end{alertblock}

\end{column}

\begin{column}{0.02\textwidth}\end{column}

\end{columns}

\begin{columns}[t]

\begin{column}{.02\textwidth}\end{column}

\begin{column}{0.47\textwidth}

\begin{block}{Future Work}

\begin{itemize}
\item Bayesian estimation of Human-Cloth relationship using \emph{Proprioceptive} information of robot.
\vspace{3mm}
\item Develop motor skills learning algorithm by incorporating real-time estimates.
\end{itemize}

\end{block}

\end{column}

\begin{column}{0.47\textwidth}

\begin{block}{References}

\nocite{*}
\scriptsize{
\bibliographystyle{unsrt}

\begin{thebibliography}{10}
\bibitem{tamei}
T. Tamei, T. Matsubara, A. Rai, and T. Shibata, ``Reinforcement learning of clothing assistance with a dual-arm robot'' in \emph{Proc. of the 11th IEEE-RAS Int. Conf. on Humanoid Robots, pp. 733-738,} 2011.

\bibitem{topology}
E. S. L. Ho and T. Komura, ``Character motion synthesis by topology coordinates,'' in \emph{Proc. of EUROGRAPHICS2009, vol. 28, no. 2, pp. 299-308,} 2009.

\bibitem{ellipse}
M. C. Shakarji, ``Least-squares fitting algorithms of the NIST algorithm testing system.'', in \emph{Journal of Research-National Institute of Standards and Technology 103,} 1998.
\end{thebibliography}}

\end{block}

\end{column}

\begin{column}{.02\textwidth}\end{column}

\end{columns}

\end{frame}

\end{document}